\documentclass[11pt]{article}
\setlength{\topmargin}{0in}
\setlength{\headheight}{0in}
\setlength{\headsep}{0in}
\setlength{\textheight}{8.7in}
\setlength{\textwidth}{6.5in}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\parindent}{0.4in}
\setlength{\parskip}{0.10in}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\usepackage{setspace}
\setstretch{2}

\begin{document}

\title{Rape in Greek Myth}
\author{Miles Wu (Preceptor: Dawn LaValle)}
\maketitle

%\begin{abstract}

%\end{abstract}

\newpage

\section{Introduction}
The motif of rape occurs frequently across Classical Greek myths, but it does not always appear under its modern guise.
To us, while the word `rape' refers to the lack of consent on the woman's part to engaging in sexual intercourse, in many Greek myths it does not seem to mean that.
For example, in the infamous myth of the Rape of Helen of Troy, many think that Helen was not  forced against her will to follow and make love to Paris, as Aphrodite has caused her to become infatuated with Paris, yet the myth is still often called a rape.

In an attempt to explain this discrepancy between the meaning of the word `rape' in modern translations of Greek myth and its current contemporary meaning, I look to the etymological root of the English word `rape', which is a Latin word, \emph{raptus} (the verb being \emph{rapio}).
This means `seized' or `taken by force' and is often used to refer to physical objects being stolen.
While it can have our modern meaning of sexual violation, this was very uncommon as another Latin word was used instead (\emph{stuprum}).
Perhaps in these Greek myths `rape' takes on \emph{raptus}' more common meaning, namely that of stealing.
This explanation implies that the women involved in these myths are treated as the property of a male and their consent is not considered at all in labeling their abduction a `rape'.

In this paper I will look at a couple of depictions of rape in Greek myth to see if this view of `rape' as stolen property is a more suitable and accurate perspective from which to look at and understand the stories, compared to the more standard view of rape as a sexual violation.
More generally I will also evaluate whether or not women in these myths are treated as objects of male ownership.
In particular I will be examining the primary sources of two famous mythological rapes, Helen of Troy and The Rape of Persephone.

\section{Helen of Troy}
%Intro about Helen of Troy?
The myth of Helen of Troy is told primarily through two epic poems written by Homer, the \emph{Odyssey} and the \emph{Iliad}.
In this myth, Helen, the wife of Menelaus, follows Paris back to Troy as his new wife after he visits on a diplomatic mission.
Enraged, Menelaus gathers up armies from various Greek city-states and leads them in war against the Trojans.

%Establish consent
Helen's main appearance in the \emph{Iliad} occurs in book three, where Paris offers to duel Menelaus in single combat to decide the fate of the whole war.
After Paris is saved from by Aphrodite, Aphrodite appears to Helen in the disguise of an old friend and tells her that ``Paris is calling for you'' \cite[book 3, line 450]{Iliad}.
After Aphrodite's short description of how Paris is ``glistening in all his beauty'' and how he is waiting ``in the bedroom'' for her, it is said that these ``the heart in Helen's breast began to race'' \cite[book 3, line 456]{Iliad}.
This makes it apparent that she is in love with Paris and freely wants to go visit him in his bedroom.
Once she reaches the bedroom, Paris ``led the way to bed'' and she ``went with him'' \cite[book 3, line 525]{Iliad}.
There is no mention of Paris compelling her to go to his bed and the use of the word `led' implies that she followed him on her volition as it is not a word with a forceful nuance.
One could make the claim that use of the word `led' was a choice that the translator made and that the original Greek word did not have the same nuances of consent.
However, in another translation of the \emph{Iliad}, for that line it says ``his wife followed him'' instead \cite[book 3, line 448]{Iliad-rieu}.
The use of the word `follow' here also has the same implication that she willingly went into bed with Paris.

%Prove previous consent
The above is only one instance where it is shown that she is not being forced by Paris, but it can be inferred from other passages that it applies generally.
For example, when Helen is talking to Priam, once again she uses the word `follow' by saying that she ``followed your son to Troy'' \cite[book 3, line 210]{Iliad}.
After Paris is defeated by Menelaus, Helen initially does not want ``to share that coward's bed once more'' and says she'll ``never go back  again'', though she later relents as discussed in the preceding paragraph \cite[book 3, line 475]{Iliad}.
The fact that she chooses not to go to make love to Paris, albeit only for a brief time, shows that she has free will in these matters and is not being forced.
Furthermore, since it says sharing his bed once more, it shows that they have had sex before and this is also supported by Paris saying that they ``loved in love on Rocky Island'' \cite[book 3, line 522]{Iliad}.
%conclude consent
As a result I conclude that Helen was actually in love with Paris and willingly consented to sexual activities.

%blame game - what do others think
Another approach of determining whether Paris forced Helen into sex is to look at the issue of blame.
There are three obvious choices for people on whom the Trojan war can be blamed: Paris, Helen and Aphrodite.
Which choice makes the most sense depends on how one interprets the myth.
For instance, if Paris forcefully abducted Helen and raped her in our modern definition of rape then for him to shoulder the blame is obvious.
If Helen deserted her husband and willingly left for Troy, having fallen in love with Paris, then she should carry some of the blame.
Finally, according to many versions of the myth, Aphrodite caused Helen to fall in love with Paris as a reward for Paris choosing her in the divine beauty contest.
If one is to believe this, then perhaps Aphrodite deserves some of the blame for meddling.
Looking at where characters in the \emph{Iliad} and \emph{Odyssey} place the blame gives clues as to which interpretation of the myth is most accurate.

When Odysseus visits the underworld in Book 11 of the \emph{Odyssey}, he meets various slain Achaean heroes and they talk about the war.
Twice he refers to Helen in a negative light and sees her actions as the root cause of the war.
The first time he bemoans to Alcinous the loss of his comrades who ``perished all for the sake of a vile woman [Helen]'' \cite[book 11, line 384]{Odyssey}.
The second time he expresses his pity at Agamemnon's misfortunes that have all come about ``through the schemes of women'' \cite[book 11, line 438]{Odyssey}.
When he speaks of `women' in the previous quote, he is referring to two women in particular, Helen and Klytaimestra (Agamemnon's wife), because in the next sentence he says that ``many of us died for the sake of Helen, and ... Klytaimestra plotted treason against you'' \cite[book 11, line 438]{Odyssey}.
It is clear from these that Odysseus places the blame of the Trojan war on Helen alone.
In book three of the \emph{Iliad}, when Helen goes up onto the walls of the city to watch the duel between Menelaus and Paris, Priam greets Helen and makes it clear that he doesn't ``blame her'' \cite[book 3, line 198]{Iliad}.
Although this is different from Odysseus, he still does not accuse Paris, but instead chooses to ``hold the gods to blame'' as ``they are the ones who brought this war'' \cite[book 3, line 200]{Iliad}.

Were Paris to have abducted and violated her against her will, everyone would surely blame him for the Trojan war.
However, in both the \emph{Iliad} and \emph{Odyssey} it is not the case that people hold Paris responsible for the war and in fact most people find Helen to be at fault.
This implies that Helen willingly left Menelaus for Paris, either with or without Aphrodite's meddling.

%Transition to property
There seems to be no evidence in either the \emph{Iliad} or the \emph{Odyssey} for Helen being forced to go unwillingly to Troy with Paris, or for her being sexually violated by Paris.
Everything points to her actually being in love with Paris, following him voluntarily to Troy and consenting to sharing his bed.
This is obviously incompatible with the modern definition of `rape' and motivates our looking for evidence that another meaning for this word is being used.
As suggested in the Introduction, there is a linguistic justification for why `rape' might have the meaning of stolen property.
In order to see if this meaning is appropriate, I look for evidence that Helen, in the context of her removal, is objectified and treated more as an item owned by Menelaus and stolen by Paris.

%Property aspect
Many people in the \emph{Iliad} when speaking of Helen's removal to Troy use language that makes it seem as if she is merely an object that was stolen, like a spoil of war.
Hector, when he is rebuking Paris for being a coward, says that Menelaus is the ``man you robbed of his sumptuous, warm wife'' \cite[book 3, line 61]{Iliad}.
Generally the word `rob' is applied to property that is taken unlawfully, so it seems odd to use it to refer to Helen who is a person who voluntarily left Menelaus for Paris unless Hector considers her to be stolen property of Menelaus.
In the negotiations for the duel between Menelaus and Paris, they also always mention Helen along with her treasures or wealth.
One example of this is Paris saying that they'll ``fight it out for Helen and all her wealth'' \cite[book 3, line 85]{Iliad}.
This makes it seem like Helen is just a belonging of Menelaus that has been stolen just like how her wealth was stolen.
Even Helen herself recognizes that she is treated as the property of some man.
When Menelaus and Paris are dueling, she knows that the duel will settle her fate regardless of what her opinion and wishes are.
This is shown when she is arguing with Aphrodite after the duel, where she says that if ``Menelaus [had] beaten your handsome Paris ... he longs to take me home''  \cite[book 3, line 466]{Iliad}.

%Culture of kidnapping/trophies
Treatment of women as property of men is not just exclusive to Helen and her `rape'; it is rife throughout many Greek myths.
The \emph{Iliad} itself even starts \emph{in medias res} with the kidnapping of Chryses' daughter by Agamemnon and the attempts of Chryses to ``win his daughter back'' \cite[book 1, line 14]{Iliad}.
During a quarrel with Achilles,  Agamemnon orders his heralds to ``go to Achilles' lodge [and] take Briseis'', Achilles' girl \cite[book 1, line 379]{Iliad}.
Briseis herself was a ``girl that he had won from Lyrnessus ... when he sacked'' it \cite[book 2, line 690]{Iliad-rieu}.
Throughout ancient times it was a standard practise for victors, after winning a battle, to plunder the possessions of the loser, but from the kidnappings of Chryses and Briseis one sees that women are treated as possessions that can also be stolen as war trophies.

%Women's views/actions not important
At first glance, it is perhaps odd to think that Helen only has a few appearances in both the \emph{Iliad} and \emph{Odyssey}, despite her being the root cause of the entire war.
Even her main appearance in the \emph{Iliad} in Book three is not that lengthy: she only talks to Priam about the Greek warriors and goes to Paris after the duel.
The reader is only sparingly told of her feelings and desires, because no-one asks Helen of what she wants, not even Paris or Menelaus or the gentle and wise Priam.
From this it can be inferred again that Helen (and more generally women in Greek myths) has been treated much less as a person and more as an item of property that has no feelings.

%evidence for wishes not respected
As established earlier she fell in love with Paris and willingly followed him to Troy, yet Menelaus would not accept her wish and went to war over it.
In the \emph{Iliad} she hints at a wish to return to Menelaus, as she wishes that Paris would have been ``brought down by that great soldier, my husband'' and she was filled with ``yearning warm and deep for her husband long ago, her city and her parents'' \cite[book 1, line 500/168]{Iliad}.
In the \emph{Odyssey}, she says that after much time in Troy her ``heart had changed by now and was for going back home again'' \cite[book 11, line 260]{Odyssey}.
Despite her wanting to go back to Greece, she still ends up staying in Troy, and beforehand, despite wanting to go to Troy, Menelaus does not accept this; Helen's wishes do not seem to be respected, much like an item without desires.

%Preceptor suggestions
%See what passages in particular talk about the removal of Helen to Troy and
%how they talk about it.
%Does Helen have an opinion about what her ``rape" means? Do the other characters like Priam or the other men of Troy in Book 3 of the \emph{Iliad}?

\section{The Rape of Persephone}
%Intro about Persephone?
The myth of the Rape of Persephone is told in the \emph{Homeric Hymn to Demeter} and is about Hades kidnapping Demeter's daughter, Persephone, to the underworld.
At first Demeter cannot find her and her despair causes the earth not to bear fruit or crops.
With people throughout the world starving in this famine, Zeus realises that in order to rectify this he must persuade Hades to return Persephone to Demeter.
Although this story has at its center a `rape' just like the Helen myth, there are both striking differences and similarities between the two.

%def abduction
Unlike Helen who went to Troy willingly, it is well beyond doubt that Persephone was forcefully abducted by Hades as there are multiple references to it being against her will in the poem.
When Hades kidnapped her, it says that ``against her will he seized her and on his golden chariot carried her away as she wailed'' \cite[line 19]{persephone}.
As if the phrase `against her will' is not clear enough, she is wailing as Hades takes her away which is clear that she is not content with what is happening to her.
To really emphasize this point, the author uses the the same phrase again when Persephone is relating to Demeter how Hades ``against my will and by force he made me taste of'' the dreaded pomegranate seed that forces her to return to the underworld every year for a third of the year \cite[line 412]{persephone}.

% but maybe not rape
However, there is again no evidence that Persephone was forced into having sexual intercourse which is our modern definition of the word `rape'.
Although the messenger Hermes, when he visits Hades, said that he found her ``sitting on his bed ... she was in many ways reluctant and missed her mother'' \cite[line 344]{persephone}, this does not necessarily imply any intercourse; there are many other benign explanations.
For instance, she could have simply been sitting on his bed and Hades, recognising her unwillingness and respecting her wishes, did not engage in any sexual activities.
Unlike with Helen and Paris where there was direct evidence from the \emph{Iliad} that they had sex, there is no such evidence with Persephone, so it cannot be said that there was a `rape' in the modern sense of the word.
%so what is it
Again, like with the Helen myth, a definition of `rape' that is more compatible with the myth needs to be found, and the notion of stolen property fits better.

%Zeus father gave away
There is again a strong indication that she has been treated as property, though this time of her parents and not of a husband.
The first hint of this is that Zeus, Persephone's father and Hades' brother, has given permission to Hades to perform the abduction: ``By Zeus' counsels, his brother ... was carrying her away'' \cite[line 30]{persephone}.
With Zeus giving permission to abduct Persephone it shows that he considers her to be his own property, which he can give away and do whatsoever he wishes with, and that her own wishes as another person and not an object are not considered at all.
In the account this is supported by Hyperionides who says that ``no other immortal is to be blamed save cloud-gathering Zeus who gave her to Hades'' \cite[line 78]{persephone}.
The other interesting to note here is the author's use of the word `gave', which lends yet further support to this argument.

% Zeus gets her back
Zeus is also the one who ends up retrieving her from the underworld, as he is the one who sends Hermes to relay the request to Hades, not because she wishes to come back, but because he wants to pacify Demeter.
This lack of consideration of her desire by Zeus again indicates that he thinks that she is his property that he can do whatever he desires with.
Hades happily complies with this request as ``with smiling brows [he] obeyed the behests of Zeus'' \cite[line 388]{persephone}.
Hades offering her back so willingly illustrates that he also thinks that she is still the property of Zeus and he does not wish to be a thief.
%evidence for parents in control
Zeus treating Persephone as his own property is again seemingly not that unusual as Demeter says as a blessing to some mortals,
``may all who dwell in the Olympian halls grant you men to wed and bear children as your parents wish'' \cite[line 136]{persephone}, implying that people getting married and bearing children are controlled by their parents' wishes.

%persephone doesn't get much of a mention either
Once again Persephone, despite being the center of the story, is not the main character of the homeric hymn, but it is her mother that is the focus.
At the core of the myth is the `rape'.
If it were the modern sense of this word, it would seem fitting that Persephone's perspective is the most important, but in the sense relating to stolen property, it would be most appropriate to select the owner's viewpoint.
This seems to be what is happening here: Demeter bemoans the sudden loss of her daughter, which she considers to be her property.
Therefore she ends up being in the spotlight instead of Persephone.

%why such outrage
One large difference between the Helen myth and the Persephone myth is the reaction of everyone else to the `rape'.
In the Helen myth all the Greek city-states unite and march forward in battle against the Trojans, but in the Persephone myth only Demeter seems to be enraged.
This can perhaps be attributed to the issue of ownership of the girls involved.
For Helen, it was undeniable that she would have been the property of Menelaus and that she was stolen by Paris, which is sure to provoke a large reaction from Menelaus and his allies.
On the other hand for Persephone, she was the property of both her parents, Zeus and Demeter.
Zeus gave her away to Hades, so her kidnapping by Hades was in fact semi-sanctioned and not outright stealing.
In a male-oriented society it is plausible that the father's opinion holds a larger weight than the mother's, and therefore no-one else besides Demeter thought it was theft and did not react badly to the kidnapping.

\section{Conclusion}
Using the modern strict definition of forced sexual intercourse, there is no evidence for either Helen or Persephone being raped.
In Helen's case it is clear that she willingly went and made love to Paris and there is no indication that any sexual activity occurred with Persephone.
From an etymological viewpoint, it can be argued the titular `rape' instead refers to the women being stolen like property and this is supported by evidence in the two myths.
Both Helen and Persephone are often treated by the other characters as objects, owned by either a husband or a parent.
Evidence for women generally being considered as owned by another is shown in both myths through war trophy girls and Demeter's blessing.
Both women also do not feature prominently in the written account, despite being the center of the myth, and their wishes and desires are not respected by their supposed owners, just like an object would not have many appearances in a story and they do not have wishes or desires.

It is also easy to see how confusion between these two definitions of `rape' could occur because when women are objectified and treated as property their wishes are not necessarily respected (as evidenced by the forceful kidnapping of Persephone); it is possible that a man ignores their wish not to have sex.
As such one definition can imply the other.
But my argument for why the property-based definition of `rape' makes more sense, other than the linguistic justification, rests upon their being no direct evidence for sexual violation in either case and there being evidence that they are treated as stolen property.
Even if Persephone had been forced to have sex by Hades, this would not invalidate the argument, because I argue the reason it was called a `rape' would not be this but instead the stolen property aspect.







\newpage
\begin{thebibliography}{9}
	\bibitem{Iliad}
        Homer. \emph{The Iliad of Homer}. Trans. Richard Lattimore. New York: HarperCollins, 2007.
	\bibitem{Iliad-rieu}
		Homer. \emph{The Iliad}. Trans. E. V. Rieu. London: Penguin Books, 2003.
	\bibitem{Odyssey}
        Homer. \emph{The Odyssey of Homer}. Trans. Richard Lattimore. Chicago: University of Chicago Press, 1961.
	\bibitem{persephone}
        \emph{The Homeric Hymns}. Trans. Apostolos N. Athanassakis. Baltimore: The John Hopkins University Press, 2004.
\end{thebibliography}

\end{document}